---
title: "Pipeline"
format: html
---

This notebook is the orchestrator level of the pipeline.
It uses `targets` and `tar_tangle` to create a data processing DAG
for the ERA5-Land data. The DAG is defined in each code chunk as a 
`tar_target` object when it is assigned, while `tar_tangle()`
is used to run the pipeline as a `targets` script.

```{r}
library(here)
library(targets)
library(tibble)
library(tidyr)
library(dplyr)
library(lubridate)
library(purrr)
library(stringr)
library(reticulate)
```

First, acquire the input data, which requires the healthsheds from
the Gold Mine. We'll use `stagecoach` for this as usual:

```{r}
#| format: file
#| packages: [here]

library(here)

stagecoach_manifest <- here("stagecoach_manifest.yml")
download_era5_script <- here("src", "download_era5.py")
```

```{r}
#| format: file
#| packages: [here]

input_files <- list.files(
    here("data", "inputs"),
    recursive=T,
    full.names=T
)
```

We'll track the configuration of the CDS API call as a separate yaml target:

```{r}
#| packages: [tibble, stringr, dplyr, lubridate, tidyr]

cds_api_config <- 
    expand_grid(
        years = 2010:2025,
        months = str_pad(1:12, 2, pad = "0"),
        days = 1:31,
        hours = str_pad(0:23, 2, pad = "0") |>
            paste0(":00"),
        countries = c("MDG", "NPL"),
        data_format = "grib",
        download_format = "unarchived",
        datasets = c("reanalysis-era5-land", "reanalysis-era5-single-levels"),
        variables = c(
            "2m_temperature",
            "2m_dewpoint_temperature",
            "volumetric_soil_water_layer_1",
            "volumetric_soil_water_layer_2",
            "surface_solar_radiation_downwards",
            "total_precipitation",
            "10m_u_component_of_wind",
            "10m_v_component_of_wind",
            "surface_pressure",
            "soil_temperature_level_1",
            "geopotential"
        )
    ) |> 
    filter(
        # filter out invalid dates
        !(months %in% c(4, 6, 9, 11) & days == 31) &
        !(months == 2 & days > 29) &
        !(months == 2 & days == 29 & !leap_year(years))
    ) |>
    filter(
        # filter out time invariant variables; only need the first hour for geopotential
        !(variables == "geopotential" & hours != "00:00")
    )
    
```

We can see one of these groups as an example:

```{r}
{
    cds_api_config |>
        sample_n(1)
}
```

This group constitutes one call to the API, which will be done in Python using
the `cdsapi` package. The next step is to write the Python script that will make 
the API calls.

To send an API call, we'll need to convert the tibble to a dictionary:

```{r}
#| packages: [purrr, reticulate]
cds_api_config_dict <- cds_api_config |>
    
    # every single combination of the above variables is too
    # granular, so we'll group by country, dataset, variable, year, and month
    group_by(countries, datasets, variables, years, months) |>
    # this leaves us with just over 816 calls to make, which is still a lot but more manageable
    group_split() |>
    map(~ {
        list(
            "data_format" = unique(.x$data_format),
            "download_format" = unique(.x$download_format),
            "dataset" = unique(.x$datasets),
            "variable" = unique(.x$variables),
            "country" = unique(.x$countries),
            "year" = unique(.x$years),
            "month" = unique(.x$months),
            "day" = unique(.x$days),
            "time" = unique(.x$hours)
        )
    })
```

The Python function defined in `download_era5.py` will take this dictionary as input and make the API call.
We'll import it to R using `reticulate`:


```{r}
library(reticulate)

source_python(download_era5_script)

download_era5(cds_api_config_dict[[6336]], here("sandbox"), dry_run = FALSE)
```

During testing, all the examples less than index 6337 worked, but the one at 
6337 threw an error. Let's double check this:

```{r}
{
    set.seed(as.numeric(now()))
    #rand_index <- sample(6337:length(cds_api_config_dict), 1)
    rand_index <- sample(1:6337, 1)
    print(rand_index)
    download_era5(cds_api_config_dict[[rand_index]], here("sandbox"), dry_run = FALSE)
}

```

This is because the API call for that particular 
combination of parameters is invalid, likely due to the fact that there is no data for that specific combination of country, dataset, variable, year, and month.
Pretty straightforward! We can use targets to run all of the API calls in
parallel, right from this notebook:

```{r}
#| pattern: map(cds_api_config_dict)
#| packages: [reticulate, here]

raw_era5_download_file <- {
    
    if(!testing){
        source_python(download_era5_script)
        res <- download_era5(cds_api_config_dict[[1]], here("data", "intermediates"), dry_run = FALSE)
        res
    } else {
        # During testing, we'll just touch the 
        print(download_era5_script)
        print(cds_api_config_dict[[1]])
        
    }
}
```