---
title: "Pipeline"
format: html
---

This notebook is the orchestrator level of the pipeline.
It uses `targets` and `tar_tangle` to create a data processing DAG
for the ERA5-Land data. The DAG is defined in each code chunk as a 
`tar_target` object when it is assigned, while `tar_tangle()`
is used to run the pipeline as a `targets` script.

```{r}
library(here)
library(targets)
library(tibble)
library(tidyr)
library(dplyr)
library(lubridate)
library(purrr)
library(stringr)
library(reticulate)
```

First, acquire the input data, which requires the healthsheds from
the Gold Mine. We'll use `stagecoach` for this as usual:

```{r}
#| format: file
#| packages: [here]

library(here)

stagecoach_manifest <- here("stagecoach_manifest.yml")
download_era5_script <- here("src", "download_era5.py")
```

```{r}
#| format: file
#| packages: [here]

input_files <- list.files(
    here("data", "inputs"),
    recursive=T,
    full.names=T
)
```

We'll track the configuration of the CDS API call as a separate yaml target:

```{r}
#| packages: [tibble, stringr, dplyr, lubridate, tidyr]

# assign temp variables
{
}

cds_api_config <- 
    expand_grid(
        years = 2010:2025,
        months = 1:12,
        days = 1:31,
        hours = str_pad(0:23, 2, pad = "0") |>
            paste0(":00"),
        countries = c("MDG", "NPL"),
        data_format = "grib",
        download_format = "unarchived",
        datasets = c("reanalysis-era5-land", "reanalysis-era5-single-levels"),
        variables = c(
            "2m_temperature",
            "2m_dewpoint_temperature",
            "volumetric_soil_water_layer_1",
            "volumetric_soil_water_layer_2",
            "surface_solar_radiation_downwards",
            "total_precipitation",
            "10m_u_component_of_wind",
            "10m_v_component_of_wind",
            "surface_pressure",
            "soil_temperature_level_1"
        )
    ) |> 
    filter(
        # filter out invalid dates
        !(months %in% c(4, 6, 9, 11) & days == 31) &
        !(months == 2 & days > 29) &
        !(months == 2 & days == 29 & !leap_year(years))
    ) |>
    # every single combination of the above variables is too
    # granular, so we'll group by country, dataset, variable, year, and month
    group_by(countries, datasets, variables, years, months) |>
    # this leaves us with just over 816 calls to make, which is still a lot but more manageable
    group_split()
    
```

We can see one of these groups as an example:

```{r}
{
    cds_api_config[[675]]
}
```

This group consititutes one call to the API, which will be done in Python using
the `cdsapi` package. The next step is to write the Python script that will make 
the API calls.

To send an API call, we'll need to convert the tibble to a dictionary:

```{r}
#| packages: [purrr, reticulate]
cds_api_config_dict <- cds_api_config |>
    map(~ {
        list(
            "format" = unique(.x$data_format),
            "download_format" = unique(.x$download_format),
            "dataset" = unique(.x$datasets),
            "variable" = unique(.x$variables),
            "country" = unique(.x$countries),
            "year" = unique(.x$years),
            "month" = unique(.x$months),
            "day" = unique(.x$days),
            "time" = unique(.x$hours)
        )
    })
```

The Python function defined in `download_era5.py` will take this dictionary as input and make the API call.
We'll import it to R using `reticulate`:


```{r}
library(reticulate)

source_python(download_era5_script)

download_era5(cds_api_config_dict[[1]], area = c(90, -180, -90, 180), output_file = "test.grib", test = TRUE)
```